### 2019-06-21 draft.tex
apply some suggestions from Tomas

\documentclass{llncs}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{listings}
\usepackage[usenames,dvipsnames]{color}
\usepackage{alltt}
\usepackage{xspace}
\usepackage{multirow}
\pagestyle{plain}
\usepackage{xcolor}
\usepackage{algorithm2e}
\usepackage{tikz,pgfplots}
\usepackage{etoolbox}
\usepackage{mathtools}
\usepackage{caption}
%\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage{listings}
%\usepackage{adjustbox}
%\usetikzlibrary{arrows,calc}
\captionsetup{compatibility=false}
%\usepackage{a4wide}
%\usepackage[margin=40mm, top=35mm]{geometry}

%\usepackage{latexsym}
\usepackage{setspace}
\usepackage{cancel}
\usepackage{graphicx}
\usepackage{appendix}
\usepackage{amssymb}
%\usepackage{times}

%here are alternatives for the times package 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage{times}
%-------
%\usepackage{pslatex}
%-------
\usepackage{mathptmx}
\usepackage{mathptm}
%-------
%\usepackage{mathtools}
%\usepackage{newtxmath} 
%\usepackage{newtxtext}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \usepackage{fullpage}

%\usepackage{cancel}
%\usepackage{verbatim}
%\usepackage{chngpage}
%\usepackage{xcolor}
%\usepackage{mathrsfs}

\lstset{language=Java,
	basicstyle=\ttfamily\footnotesize,
	showspaces=false,
	showtabs=false,
	breaklines=true,
	showstringspaces=false,
	breakatwhitespace=true,
	commentstyle=\color{pgreen},
	keywordstyle=\color{blue},
	stringstyle=\color{red},
	basicstyle=\ttfamily
}

\newcommand{\Var}{\mathtt{Var}}
\newcommand{\Exp}{\mathtt{Exp}}
\newcommand{\Cmd}{\mathtt{Cmd}}
\newcommand{\Grd}{\mathtt{Grd}}
\newcommand{\Prg}{\mathtt{Prg}}
\DontPrintSemicolon
\SetKwRepeat{Loop}{Loop\{}{\}}%

\newcommand{\true}{\mathsf{true}}
\newcommand{\false}{\mathsf{false}}

\newcommand{\todo}[1]{{\color{blue}TODO:#1}}
\newcommand{\hide}[1]{}
\newcommand{\defn}{\overset{\triangle}{=}}
\newcommand{\maxi}{m}
\newcommand{\cur}{cur()}
\newcommand{\dom}[1]{\mathsf{dom}(#1)}
\newcommand{\ite}[3]{
	 \ifmmode
	 \mathbf{if}\ #1 \ \mathbf{then}\ #2\  \mathbf{else}\ #3
	 \else
	 \textbf{if}\ #1 \ \textbf{then}\ #2\  \textbf{else}\ #3
	 \fi}
\newcommand{\rloop}{
	\ifmmode
	\mathbf{Loop}
	\else
	\textbf{Loop}
	\fi}

\newcommand{\Z}{\mathbb{Z}}

\title{\hspace{-0.1cm}{J-ReCoVer}:~Java~Reducer~Commutativity~Verifier}
\author{}
\institute{}
\hide{
\author{
Yu-Fang Chen\inst{1}\inst{2}
\and
Chang-Yi Chiang\inst{2}
\and
Hsin-Hung Lin\inst{1}
\and
Wei-Tsung Kao\inst{1}
\and
Yean-Fu Wen\inst{2}
\and
Wei-Cheng Wu\inst{1}
}
\institute
{
Institute of Information Science, Academia Sinica, Taiwan
\and
Graduate Institute of Information Management, National Taipei University, Taiwan
}}
\author{}
\institute{}

\begin{document}

\maketitle

\begin{abstract}

MapReduce framework for data-parallel computation was first proposed by Google~\cite{dean04} and later implemented in the Apache Hadoop project~\cite{hadoop}.
Under the MapReduce framework, a reducer computes output values from a sequence of input values transmitted over the network.  Due to the non-determinism in data transmission, the order of input values arrives at the reducer is not fixed.
The \emph{commutativity problem} of reducers asks if the output of a reducer is independent of the order of its inputs. There are several advantages for a reducer being commutative, e.g., the verification problem of a MapReduce program can be reduced to the problem of verifying a sequential program.
We present a tool J-ReCoVer (Java Reducer Commutativity Verifier) that implements effective heuristics for reducer commutativity analysis. J-ReCoVer is the first tool that is specialized in checking reducer commutativity. Experimental results over 118 benchmark examples collected from open repositories are very positive; J-ReCoVer correctly handles over 97\% of them.

\end{abstract}

\section{Introduction}
\label{section:introduction}

MapReduce is one of the most popular frameworks for data parallel computation.
A MapReduce~\cite{dean04,hadoop} program consists of several pairs of \emph{mappers} and \emph{reducers} running on a machine cluster for handling big data in parallel. Usually, mappers and reducers are the only components in a MapReduce program that involves concurrency. Mappers read data from a distributed database and output a sequence of \emph{key-value} pairs. The sequence elements (key-value pairs) with the same key will be sent to the same reducer for further processing. Due to scheduling policies and network latency, the same inputs may arrive at a reducer in a different order in different executions, which results in the so-called \emph{commutativity problem} of reducers~\cite{csallner13testing,xiao14mr,ChenHSW15,ChenSW16}, that is, if the output of a reducer is independent of the order of its inputs.
If all reducers are commutative, they will have the same external behavior under all possible schedules. In such a case, by fixing a schedule, the verification problem of a MapReduce program can be reduced to the verification problem of a sequential program, which is known to be much easier than the verification problem of concurrent programs. One needs to consider only a particular interleaving of input values instead of all of them.

\hide{More concretely, for the verification of a MapReduce program, we propose to do it in two phases. First, ensuring if all reducers are commutative. If some reducers are non-commutative, modifying them to commutative ones. Usually, the modification is not a difficult task~\cite{xiao14mr}, and can be done without affecting their functionality and performance. For example, assume that the task of a reducer is to find the name of the person with highest score. Such a reducer is non-commutative when the input include two people with the same highest score. This reducer can be made commutative, by also comparing the ID number of people with the same score. Once all reducers are made commutative, we reduce the verification problem to sequential verification by fixing a scheduler. In this two-phase approach, the key enabling technique is an efficient procedure for checking reducer commutativity.}

On the other hand, the non-commutative behavior of a reducer is often the source of very tricky bugs. A study conducted by Microsoft investigated the commutativity problem of $508$ reducers running on their MapReduce server~\cite{xiao14mr}. The reducer programs have been checked carefully using all traditional means such as code review, testing, and experiments with real data for months. Still, among them, five programs contain very subtle bugs caused by non-commutativity (confirmed by the programmers).

However, checking reducer commutativity is a difficult problem on its own right~\cite{ChenHSW15,ChenSW16,ChenLTW17}. Even for a simple case that all values are mathematical integers, it is proved undecidable in~\cite{ChenHSW15}. For the case that all values are machine integer (e.g, 64-bits integer), the problem is decidable, but the only available algorithm that was proposed in the same paper is of very high complexity and hence is only of theoretical interests.

In this paper, we propose heuristics for the commutativity problem that works very efficiently on a large set of practical integer reducer programs. We reduce the commutativity problem to an SMT problem in a sound but incomplete manner. We also propose optimizations to enable the approach that works on real-world examples. For the case that the reducer is not proven commutative, we complement the approach by using testing to find concrete counterexamples. We implement all these heuristics as a tool called J-ReCoVer (Java Reducer Commutativity Verifier), which is available at~\verb|http://jrecover.iis.sinica.edu.tw|.
We collected benchmarks from open repositories such as Github and Bitbucket to evaluate J-ReCoVer. With the help of a search engine \emph{searchcode.com} over those repositories, we collected 118 examples. J-ReCoVer is able to correctly analyze all but 3 of them.

\subsection*{Related Works}
Reducer commutativity problem can be reduced to a program equivalence problem. One creates another program $R'$ that first non-deterministically swaps two consecutive input values and then executes the code of $R$. If $R'$ and $R$ are equivalent, using the fact that all permutations of a list can be obtained by swapping consecutive list elements finitely many times\footnote{Here is an example to produce [3;2;5;1;4] from [1;2;3;4;5] by swapping consecutive elements: [\textbf{1};\textbf{2};3;4;5] $\rightarrow$ [2;\textbf{1};\textbf{3};4;5] $\rightarrow$ [2;3;1;\textbf{4};\textbf{5}]$\rightarrow$ [2;3;\textbf{1};\textbf{5};4]$\rightarrow$ [\textbf{2};\textbf{3};5;1;4]$\rightarrow$ [3;2;5;1;4].}, $R$ can be proved to be commutative. There is a series of research results on program equivalence checking (or closely related topics such as regression verification and translation validation), to name a few~\cite{Pnueli:1998:TV,fedyukovich2015automated,barthe2011relational,KlebanovRuemmerUlbrich2017}. At a high level, checking equivalence of two programs $P$ and $P'$ can be reduced to a sequential verification problem by executing $P'$ after $P$ and at the end checking if they always have the same outputs. It can be made more efficient by finding the right~\emph{synchronization points} and combining the code of $P$ and $P'$ in an interleaved manner. A lot of research effort in this direction is in finding good synchronization points. For the case of commutativity analysis, one can simply use the beginning of the loop as the synchronization point. However, in our experience, if we naively reduce the commutativity problem to equivalence problem and check it in a precise manner, many examples cannot be solved. In this paper, we propose a good over-approximation (c.f. Section~\ref{sec:prover}) of the reducer's behavior. The approximation allows a much more efficient yet precise enough commutativity analysis.

One main motivation of commutativity analysis is to reduce the verification problem of a concurrent program to a sequential one. There are several pieces of research sharing the similar motivation, e.g., the concept of \emph{robustness} of event-driven asynchronous programs~\cite{ahmed2017:robustness} and program running under relaxed memory model~\cite{ahmed2013:robustness,AbdullaACLR13,AbdullaACLR12}. However, their computation models are quite different from that of reducers and hence the result cannot be directly applied. Besides verification, another interesting research direction is the synthesis of MapReduce programs~\cite{SmithA16}. Commutativity analysis is used as a component under the synthesis framework.

\section{Notations and Definitions}
\label{section:integer-reducers}
We use $[n,m]$ to denote the set of integers $\{k\mid n \leq k\leq m\}$ and lift the equality $=$ operation to tuples in the standard way.

\hide{
\begin{equation*}
\begin{array}{rclcl}
\multicolumn{3}{l}{v\in \Var\mbox{, and }c \in \Z}&&\textmd{Variable and Constant}\\
e \in \Exp    &\ \ \  {\triangleq}\ \ \   & v \mid c \mid \cur \mid e + e \mid e - e \mid e \times e \mid *\mid\ldots & &\textmd{Integer Expression}\\
g\in \Grd  & {\triangleq}  & e < e \mid not\ g \mid g \wedge g  \mid \ldots& & \textmd{Guard}\\
s\in \Cmd  & {\triangleq}  & v:= e \mid out(v) \mid \ite{g}{s}{s} \mid s;s && \textmd{Command}\\
p\in \Prg  & {\triangleq}  & s;\rloop\{s\};s&& \textmd{Reducer Program}\\
\end{array}
\end{equation*}}
\begin{wrapfigure}{r}{3cm}
		\vspace{-0.8cm}
	\hide{
		\begin{minipage}{0.32\textwidth}
			\begin{algorithm}[H]
				$m := \cur$; \;
				\Loop{}{
					$t:=\cur$;\;
					\uIf{ $t > m$}{
						$m := t$
					}
				};
				$out(m)$\;
			\end{algorithm}
			\caption*{(a) max}
		\end{minipage}
		\hspace{-0.5cm}
	}
	\begin{minipage}{0.3\textwidth}
		\begin{algorithm}[H]
			$s := 0;c:=0$; \;
			\Loop{}{
				$s := s+\cur$;\;
				$c := c+1$
			};
			$o := s/c$;\;
			$out(o)$\;
		\end{algorithm}
	\end{minipage}
	\hide{
		\begin{minipage}{0.4\textwidth}
			\begin{algorithm}[H]
				$a:=*;s := 0; c:=0$;\;
				\Loop{}{
					$s := s+(\cur-a)^2$; \;
					$c := c+1$
				};
				$o := s/c$;\;
				$out(o)$
			\end{algorithm}
			\caption*{(c) standard deviation}
		\end{minipage}
	}
	\caption{Average}
	\vspace{-0.8cm}
	\label{fig:reducer_example}
\end{wrapfigure}
For presentation purpose, we describe a highly simplified language to characterize the essence of reducers. The model allows to describe the computation of reducers and investigate their commutativity.
Let $\Var$ be the set of variables. An \emph{integer expression} in $\Exp$ can be either a variable, constant value, a call to the $\cur$ function that reads and consumes an input value of the reducer, their combinations over basic arithmetic operations, or a non-deterministically chosen integer value $*$.
A \emph{command} in $\Cmd$ can be an assignment, a branch statement, concatenation of a sequence of commands, or $out(v)$ that outputs the value of $v$.
A \emph{reducer program} is defined as $s_1;\rloop\{s_2\};s_3$, where $s_1,s_2,s_3 \in \Cmd$, and $s_2$ is assumed to have at most one $\cur$ expression in each \emph{branch}\footnote{We define a \emph{branch} of a command $s$ as a sequence obtained by replacing every branch statement $\ite{g}{s_1}{s_2}$ in $s$ with either $(g); s_1$ or $(not\ g) ;s_2$. For example, the loop body of Figure~\ref{fig:reducer_example}(a) has two branches $t:=\cur; (t>0); m:=t$ and $t:=\cur;(not\ t>0)$ }. According to our observation over hundreds  of reducer programs in open repositories, reducer programs are almost always in this form. The $\rloop\{s_2\}$ statement enters the loop body to execute $s_2$ repeatedly until the entire input list is consumed. An example is in Figure~\ref{fig:reducer_example}.

Some reducer uses two loops to compute the output. For example, for calculating standard deviation, one first computes the average of inputs and then uses it to compute the standard deviation. In that case, we can still verify the reducer by first partitioning it into two reducers and then verifying them separately. One reducer computes the average and the other reducer assigns a non-deterministic value to the variable $a$ that is intended to store the computed average ($a := *$). The variable $a$ will be used in the computation of standard deivation. The second reducer ``over-approximates''  all possible outputs of the standard deviation computation; it computes an output even if $a$ is not the average of inputs. However, this reducer is still commutative and its commutativity verification can still be done.

\section{Overview of the J-ReCoVer tool}
\label{sec:overview}
%Hadoop, Java;
The input of J-ReCoVer is a reducer program written in Java, which is the most popular programming language used in the Hadoop~\cite{hadoop} MapReduce framework.
The J-ReCoVer tool has three main components, the \emph{Preprocessor}, the \emph{Prover}, and the \emph{BugFinder}. As the name suggests, the Preprocessor reads as input a reducer program and performs the required preprocessing.  The goal of Prover is to show that a given reducer is commutative and the goal of BugFinder is the opposite. The architecture of J-ReCoVer can be found in Figure~\ref{fig:overview-jrecover}. The user can input a reducer program to J-ReCoVer either through our web-interface or binary application installed on their own machine.

\textbf{The Preprocessor component} first compiles a reducer program to bytecode and use the tool Soot to further convert it to the so-called ``Jimple'' format, which is an intermediate language that is designed to simplify the analysis of Java programs. Under the Hadoop MapReduce framework, the permutation of the input is handled by the scheduler/shuffler component and is affected by things like network latency, which is not controllable by programmers. In order to deal with this issue, we wrote our own dummy Hadoop environment for the reducer as a part of the Preprocessor component so that the input order of the reducer is now controlled by J-ReCoVer. Finally, we perform in the Preprocessor some program transformation tasks to simplify the analysis. For example, we convert all output commands $out(e)$ inside loops to a special assignment (will be explained in Section~\ref{sec:program_trans1} and~\ref{sec:program_trans2}).

\begin{wrapfigure}{r}{7cm}
	\vspace{-0.5cm}
\centering
	\scalebox{0.7}{
		\begin{tikzpicture}

		%J-ReCoVer
		\node[draw=red!30, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=8cm,fill=red!10,minimum height=8cm] (jrecover) at (5,0.8) {};
		\node at (5,0.4) {\textbf{J-ReCoVer}};

		%JAVA reducer
		\node[draw=yellow!80, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=0.6cm,fill=yellow!20,minimum height=4cm] (reducer) at (0.5,-1.6) {};
		\node[rotate=270, align = center] at (0.5,-3.6) {Java Reducer};

		%JPreprocessor
		\node[draw=blue!50, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=3cm,fill=blue!20,minimum height=6.cm] (preprocessor) at (3,-.8) {};
		\node[] at (3,-1.2) {Preprocessor};

		%JPreprocessor->Interface
		\node[draw=green!50, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=0.3cm,fill=green!10,minimum height=1.5cm] (interface1) at (1.8,-1.6) {};
		\node[rotate=270, align = center] at (1.8,-2.4) {Web};

		\node[draw=green!50, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=0.3cm,fill=green!10,minimum height=3cm] (interace2) at (1.8,-3.6) {};
		\node[rotate=270, align = center] at (1.8,-5) {Installed app.};

		%JPreprocessor->Technologies
		\node[draw=green!50, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=2cm,fill=green!10,minimum height=1cm] (soot) at (3.2,-2) {};
		\node at (3.2,-2.6) {Soot};

		\node[draw=green!50, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=2cm,fill=green!10,minimum height=1.4cm] (symex) at (3.2,-3.3) {};
		\node at (3.2,-4) {
			\begin{minipage}[t]{0.222\textwidth}
			\centering
			Dummy\\
			Hadoop\\
			Environment
			\end{minipage}		
			};

			\node[draw=green!50, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=2cm,fill=green!10,minimum height=1.2cm] (prog_trans) at (3.2,-5) {};
			\node at (3.2,-5.6) {
				\begin{minipage}[t]{0.222\textwidth}
				\centering
				Program\\
				Transform.
				\end{minipage}
			};

		%JProver
		\node[draw=blue!50, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=3cm,fill=blue!20,minimum height=5.5cm] (prover) at (6.5,0) {};
		\node[] at (6.5,-0.4) {Prover};

		%JProver->Technologies
		\node[draw=green!50, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=2cm,fill=green!10,minimum height=1.2cm] (z3) at (6.5,-1) {};

		\node at (6.5,-1.6) {
			\begin{minipage}[t]{0.222\textwidth}
			\centering
			Z3 SMT \\
			Solver
			\end{minipage}
		};

		\node[draw=green!50, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=2cm,fill=green!10,minimum height=1.2cm] (z3) at (6.5,-2.5) {};

		\node at (6.5,-3.1) {
			\begin{minipage}[t]{0.222\textwidth}
			\centering
			Equation~\ref{eq:commu}
			\end{minipage}
		};

		\node[draw=green!50, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=2cm,fill=green!10,minimum height=1.2cm] (z3) at (6.5,-4.) {};

		\node at (6.5,-4.6) {
			\begin{minipage}[t]{0.222\textwidth}
			\centering
			Live Variable \\
			Analysis
			\end{minipage}
		};

		%JBugFinder
		\node[draw=blue!50, inner sep=0, anchor=north,rounded corners, fill opacity=0.85,text width=3cm,fill=blue!20,minimum height=0.5cm] (learner) at (6.5,-6.2) {};
		\node[] at (6.5,-6.5) {BugFinder};

		%JResults

		\node[rotate=270, align = center] at (9.5,-1) {
			Commutative
		};

		\node[rotate=270, align = center] at (9.5,-3.5) {
			Unknown
		};

		\node[rotate=270, align = center] at (9.6,-6) {
			\begin{minipage}[t]{0.25\textwidth}
			\centering
			Non-Commutative	 \\
			+ Counterexample
			\end{minipage}

		};

		\draw [->] (8,-6.6) -- (9.2,-6.6) ;
		\draw [->] (8,-1) -- (9.2,-1) ;
		\draw [-] (8,-6.3) -- (8.5,-6.3) --(8.5,-3.5);
		\draw [->] (8,-3.5) -- (9.2,-3.5) ;

		\draw [->] (4.5,-3) -- (5,-3) ;
		\draw [->] (4.5,-6.5) -- (5,-6.5) ;

		\draw [->] (0.8,-4) -- (1.5,-4) ;

		\end{tikzpicture}
	}
		\vspace{-1cm}
	\caption{Overview of the J-ReCoVer Tool. }\label{fig:overview-jrecover}
		\vspace{-0.8cm}

\end{wrapfigure}

\textbf{The BugFinder component} randomly generated 1500 pairs of lists. The lists in a pair are permutations of each other. A concrete counterexample is reported if inputting any pair of lists from the 1500 generated ones to the reducer generates different results.
Our procedure for generating random pairs is quite naive. We use five different input list of lengths 5, 7, 9, 11, and 13. For each length, we generate 300 lists and pick uniformly at random one of its permutations. Although the approach is simple, in practice it finds counterexamples in all of our non-commutative benchmarks.

\textbf{The Prover component} reduces the commutativity problem to an SMT formula satisfiability problem. At a high level, we check equivalence between a reducer program and its variant that has two consecutive inputs swapped. We show that this can be reduced to an SMT problem and give it to the SMT solver Z3 for solving. In case that Z3 proved the formula unsatisfiable, we know that swapping any two consecutive inputs of the reducer will not change its output. Because all permutations of a list can be obtained by swapping consecutive elements finitely many times, it follows that the reducer outputs the same value for all permutations of the same list of inputs.
In this case,  J-ReCoVer stops and reports that the reducer is commutative.

\section{The Preprocessor and the Prover}
\label{sec:preprocessor_prover}

Before entering the Prover component, the Preprocessor first performs a few program transformation tasks to simplify the verification task.
Assume a reducer $s_1;\rloop\{s_2\};s_3$ is given. The section begins with an introduction to the first program transformation task (Section~\ref{sec:program_trans1}) that handles the case when the loop body $s_2$ contains $out$ commands. The output of the first transformation task is a commutativity-equivalent reducer program where $out$ does not occur in $s_2$.  Then the algorithm of the Prover is explained in Section~\ref{sec:prover}. This algorithm works only in a special case where no read from input occurs in $s_1$. For general reducer programs, we introduce the second program transformation task (Section~\ref{sec:program_trans2}) whose output is a reducer program that falls in the special case required by the Prover and ``over-approximates'' all input/output behaviors of the original reducer. The commutativity of the produced reducer implies the commutativity of the original one, but not vice verse.

\subsection{The First Program Transformation Task in the Preprocessor}
\label{sec:program_trans1}
Recall that the reducer program can perform the ouput command $out(e)$ inside the loop body $s_2$. In this case, we remember the output value in a fresh variable $v$ and replace the $out(e)$ command with an assignment $v:=e$.
Notice that the value of $v$ stays the same from the point it has been assigned a value.
If the values of $v$ are the same for all permutations of input lists at the end of $s_2$, we know that the reducer is commutative w.r.t. to all output values produced inside the loop body $s_2$.

\subsection{The Prover: Reduce Commutativity Checking to SMT Solving}
\label{sec:prover}
Below we begin with a simple reducer where $s_1$ never calls the $\cur$ function. Later we will explain how to handle the general case using program transformation. The command $s_2$ can be viewed as a function $F$ that reads the values of all variables and the current input before executing $s_2$ and outputs the values of all variables after $s_2$. Formally, the function $F(n,x_1,x_2,\ldots,x_k): \Z^{k+1} \rightarrow \Z^k$ returns a tuple of values $x_1',x_2',\ldots,x_k'$, where $n$ is the current input value of the reducer, $x_i$ and $x_i'$ are the values of variables before and after the execution of $s_2$, respectively, for $i\in[1,k]$. The construction of $F$ from $s_2$ can be done in the standard way. We check if the formula below is valid for all possible values of $n_1,n_2, x_1,x_2,\ldots,x_k$.
\begin{equation}
 F(n_1, F(n_2,x_1,x_2,\ldots,x_k)) = F(n_2, F(n_1,x_1,x_2,\ldots,x_k) )
\label{eq:commu}
\end{equation}

Using the reducer that computes average (Figure~\ref{fig:reducer_example}) as an example. It has two variables $s$ and $c$. It is not difficult to compute that $F_{average}(n,s,c)=(s+n, c+1)$. The corresponding Equation~(\ref{eq:commu}) is valid because $F_{average}(n_1, F_{average}(n_2,s,c)) =F_{average}(n_1, s+n_2, c+1)= (s+n_1+n_2,c+2)=F_{average}(n_2, s+n_1, c+1)=F_{average}(n_2, F_{average}(n_1,s,c))$.

The validity of Equation~(\ref{eq:commu}) implies that swapping any two consecutive input values will not change the final variable valuation. Informally, it says that starting from the same initial valuations of variables and with two different input orders, $[n_1;n_2]$ and $[n_2;n_1]$, the values of all program variables are the same after we execute the loop body twice. The first execution reads $n_1$ and then reads $n_2$. The other execution reads the two inputs in the reverse order.

\hide{

In the example of ``average'', the values of $s$ and $c$ are the same no matter the reducer first reads $n_1$ and then $n_2$ or reads the two values in the swapped order.
So it computes the same value for the input [1;2;\textbf{3};\textbf{4};5] and any other input with two consecutive values being swapped, e.g., [1;2;4;3;5].
The swap operation can be applied arbitrarily many times to produce all possible permutations of the input.
It follows that that starting from the same variable valuation, the loop ends with the same valuation for any permutation of input values, e.g., the reducer will output the same value for [1;2;3;4;5], [1;2;4;3;5], [1;4;2;3;5], [4;1;2;3;5], etc.
Recall that the first program transformation task stores all output values in variables that will stay unchanged until the end of the loop. The validity of Equation~(\ref{eq:commu}) also implies that the execution of the loop produces the same sequence of outputs.

We explain why that the validity of Equation~(\ref{eq:commu}) is sufficient to prove the reducer $s_1;\rloop\{s_2\};s_3$ commutative. Notice that $s_1$ does not read from the input (assumed at the beginning of this subsection). The execution of $s_1$ is deterministic in the sense that from the same initial valuation, it outputs the same sequence of values and ends at the same valuation. For the execution of $\rloop\{s_2\}$, from the validity of Equation~(\ref{eq:commu}), any permutation of input values leads the reducer to the same final valuation and the same sequence of outputs at the end of the loop.
Before executing the commands $s_3$, all inputs have been read (the loop exit condition) already. Similar to the case of $s_1$, $s_3$ does not read from the input and hence the execution of $s_3$ is deterministic. It follows that for all permutations of inputs, the reducer outputs the same sequence of values.}

In the next part, we explain how to generalize our approach to support general reducers (without the restriction on $s_1$) using program transformation.

\subsection{The Second Program Transformation Task in the Preprocessor}
\label{sec:program_trans2}

\begin{wrapfigure}{r}{3.5cm}
	\vspace{-1cm}
	\begin{algorithm}[H]
	\;\;$m := \cur + 10$; \;
	\Loop{}{
		$t:=\cur$\;
		\uIf{ $t> m$}{
			$m := t$ \;
		}
	};
	$out(m)$\;\;
\end{algorithm}
	\vspace{-0.8cm}
\caption{Reducer max$^+$}
	\vspace{-0.8cm}
	\label{fig:reducer_max1}
\end{wrapfigure}

\hide{	
\begin{figure}[t]
	\begin{minipage}{0.4\textwidth}
		\begin{algorithm}[H]
			\;\;$m := \cur + 10$; \;
			\Loop{}{
				$t:=\cur$\;
				\uIf{ $t> m$}{
					$m := t$ \;
				}
			};
			$out(m)$\;\;
		\end{algorithm}
		\caption*{(a) Reducer max$^+$}
	\end{minipage}
		\begin{minipage}{0.6\textwidth}
			\begin{algorithm}[H]
				$s:=*;$\;
				\Loop{}{
					\uIf{$s=1$}{$m := \cur + 10; s:= 2$}
					\uElse{
						$t:=\cur$\;
						\uIf{ $t > m$}{
							$m := t$ \;
						}
					}
				};
				$out(m)$
			\end{algorithm}
			\caption*{(b) Reducer max$^{+\mathtt{fix}}$}
		\end{minipage}
	\caption{Reducer examples}
	\label{fig:reducer_max}
\end{figure}
}

In real-world examples, it is often that the $s_1$ part of the reducer also reads from the input. In the example in Figure~\ref{fig:reducer_max1}, the reducer program remembers the first input value in the variable $m$, increases its value by 10, and then updates its value to bigger ones if any occurs in the loop. The corresponding Equation~(\ref{eq:commu}) is valid in this case. However, the reducer is not commutative. A counterexample can be easily found. With the list $[1,2,3,4,5]$ and its permutation $[5,4,3,2,1]$ as the inputs, the reducer respectively outputs $11$ and $15$.

\begin{wrapfigure}{r}{5.3cm}
	\vspace{-1cm}
	\begin{algorithm}[H]
		$s:=1;$\;
		\Loop{}{
			\uIf{$s=1$}{$m := \cur + 10; s:= 2$}
			\uElse{
				$t:=\cur$\;
				\uIf{ $t > m$}{
					$m := t$ \;
				}
			}
		};
		$out(m)$
	\end{algorithm}	
	\vspace{-0.4cm}
	\caption{Reducer max$^{+\mathtt{fix}}$}
	\vspace{-0.4cm}
	\label{fig:reducer_max2}
\end{wrapfigure}
We handle the problem again by program transformation. To deal with the example in Figure~\ref{fig:reducer_max1}, we move the prefix $s_1$ into the loop body and use a new variable $s$ to force that the execution of $s_1$ is always before the original loop body. The result after the transformation is demonstrated in Figure~\ref{fig:reducer_max2}. The new reducer program over-approximates all possible inputs/outputs of the original one. When $s=1$, its output sequence is the same as the original reducer. When $s\neq 1$, it may produce outputs that are not possible by the original reducer.
It also satisfies all assumptions required for the Prover component:
(I) the commands before entering the loop does not have any $\cur$ expression,
(II) each branch (recall that ``branch'' is formally defined in Section~\ref{section:integer-reducers}) of the loop body has at most one $\cur$, and
(III) $out(v)$ command does not occur inside the loop body\footnote{We need to apply the first program transformation again to fulfill this assumption.}. So we can apply Equation~(\ref{eq:commu}) to prove its commutativity.
Since the new reducer has all behaviors of the original one, the new reducer being commutative implies that the original one is also commutative.

\hide{
\begin{wrapfigure}{r}{9cm}
	\centering
	\begin{minipage}{0.8\textwidth}

	\begin{algorithm}[H]
		$s'_0;s:=*;$\;
		\Loop{}{
			\lIf{$s=1$}{ $x_{i_1} := \cur; s'_1; s:=2$}
			\lElseIf{ $s =2$}{ $x_{i_2} := \cur; s'_2; s:=3$}
			\lElseIf{ $s = 3$}{$\ldots$}
			\lElseIf{ $s = m$}{$x_{i_m} := \cur; s'_m; s:=m+1$}
			\lElse{ $s_2$}
		};$s_3$\;
	\end{algorithm}
	\end{minipage}
	\caption{The second transformation task}
	\label{fig:general_program_transformation}
\end{wrapfigure}}

This approach can be easily generalize to the situation where $\cur$ occurs multiple times in $s_1$. We first rewrite the $s_1$ part of the reducer $s_1;\rloop\{s_2\};s_3$ to the form
$$s'_0;x_{i_1}{:=} \cur;s'_1.\ldots x_{i_{m}}{:=}\cur;s'_m$$, where $\cur$ never occur in $s'_0,s'_1,\ldots, s'_m$. Then uses a branch statement with garud $s=j$ to handle the case $x_{i_j} := \cur; s'_j; s:=s+1$, for all $j \in [1,m]$.

\subsection{Optimization: Live Variable Analysis}
\label{section:optimizations}
In this section, we explain how data flow analysis is used in J-ReCoVer to improve the precision of commutativity checking and the algorithm we used to construct the $F$ function in Equation~(\ref{eq:commu}).

We find that in many cases, Equation~(\ref{eq:commu}) is a condition that is too strong for the reducer to be commutative. For instance, consider a simple example. In the loop body, the input is first stored in a variable $t$ and then added to $s$. After the loop, output the value of $s$.
In this case, the $F$ function returns the updated values of both $s$ and $t$ after the execution of the loop body. Observe that
$F(c_1, F(c_2,s,t)) = (c_1+c_2, c_1)$ and $F(c_2, F(c_1,s,t)) = (c_1+c_2, c_2)$. It follows that the corresponding Equation~(\ref{eq:commu}) is invalid. The second component of the returned tuples corresponds to the value of $t$ after executing the loop body twice. Their values are $c_1$ and $c_2$, respectively, in $F(c_1, F(c_2,s,t))$ and $F(c_2, F(c_1,s,t))$. However, in this case, the value of $t$ will not affect the output of this reducer.

\hide{
\begin{figure}[hbt]
	\begin{minipage}{0.4\textwidth}
		\begin{algorithm}[H]
			$s:= 0$; \;
			\Loop{}{
				$t:=\cur$;
				$s:= s + t$\;
			};
			$out(s)$\;\;\;
		\end{algorithm}
		\caption{A commutative reducer with an invalid Equation~(\ref{eq:commu}).}
		\label{fig:reducer_opt}
	\end{minipage}
	\ \ \ \ \ \ \ \
	\begin{minipage}{0.5\textwidth}
		\LinesNumbered
		\centering
		\begin{minipage}{0.75\textwidth}
		\begin{algorithm}[H]
			$c_1;c_2;\ldots;c_i$\;
			\lIf{$g_1$}{$\ldots$}
			\lElseIf{ $g_2$}{$\ldots$}
			\lElseIf{ $g_3$}{$\ldots$}
			\lElseIf{ $g_m$}{$c_{i+1};\ldots ;c_n; out(v);\ldots$}
			\lElse{ $\ldots$}
		\end{algorithm}
	\end{minipage}
		\caption{The $s_3$ command with an $out(v)$ command in nested branches.}
		\label{fig:nested_out}
	\end{minipage}
\end{figure}

To handle such a situation, we need an algorithm to collect a set of variables whose value may affect the output.
Recall that the first program transformation task uses a set of variables $V_2=\{v_{out_k} \mid k \in [1,l]\}$, where $l$ is the number of output commands occurred in $s_2$, to collect output values occurred in the loop body $s_2$.
The set $V_2$ is exactly the set of variables that will affect the output inside the loop.}

We perform a simple backward live variable analysis to collect all variables whose value may propagate to the output command after the loop execution. In our evaluation, the ratio of examples that our approach can successfully analyze is significantly increased from $6.8\%$ to $97.5\%$. The details of the evaluation can be found in Section~\ref{section:exp}.

\hide{
Without loss of generality, we assume there is only one output command after the loop. Then the code after the loop (i.e., the $s_3$ part of the reducer) can only be in the form of Figure~\ref{fig:nested_out}.
We unfold all concatenation of commands, so now the commands $c_1\ldots c_n$ are either assignment commands $w:=e$ or branch command $\ite{g}{c'_1}{c'_2}$, and the only output command stays in the $m$-th layer of the branch command after $c_i$. For the case that $m=0$, the entire branch command in Lines 2-6 of Figure~\ref{fig:nested_out} is replaced with $out(v)$.

The data flow analysis algorithm works as follows.
A set $V_3$ whose value is initially a singleton set $\{v\}$ is used to remember the variables that might flow to the output command $out(v)$ in $s_3$ (in Figure~\ref{fig:nested_out}).
The commands and guards will be inspected in the following order $c_{n}$,$\ldots$,$c_{i+1}$, $g_m$, $g_{m-1}$, $\ldots$, $g_1$, $c_i$, $\ldots$, $c_2$, $c_1$. For the case that a guard $g_j$ for $j\in [1,m]$ is being processed, we simply add all variables occurred in $g_j$ to $V_3$.
For the case that a command $c_j$ for $j\in [1,n]$ is being processed, we need to handle two different cases:
\begin{itemize}
	\item $c_j$ is an assignment $w:=e$ and $w\in V_3$: all variables in $e$ are added to $V_3$.
	\item $c_j$ is a branch $\ite{g}{c'_1}{c'_2}$: the algorithm starts from the same set $V_3$ and processes $c'_1$ and $c'_2$ individually and independently (in a similar manner) to obtain $V_3^1$ and $V_3^2$. After that, the algorithm takes their union $V_3^1\cup V_3^2$ as the new value of $V_3$.
\end{itemize}
After the entire $s_3$ is being processed we get the set of important variables that might affect the outputs that happen in $s_3$.
Then we modify the Equation~(\ref{eq:commu}) to the one below
\begin{equation}
F(n_1, F(n_2,x_1,x_2,\ldots,x_k) )=_{PosF(V_2\cup V_3)} F(n_2, F(n_1,x_1,x_2,\ldots,x_k) )
\label{eq:commu2}
\end{equation}
The function $PosF(X)$ returns the corresponding positions of variables in $X$ in the output of $F(n,x_1,\ldots,x_k) $. E.g., if $X = \{x_1,x_4,y\}$, where $y\notin \{x_1,\ldots,x_k\}$, then $PosF(X)=\{1,4\}$.
The use of Equation~(\ref{eq:commu2}) is critical to the success of our approach. }
\hide{
\subsection{Computing the $F$ function using symbolic execution with an efficient data structure}
\label{section:opt2}

The computation of $F$ in Equation~(\ref{eq:commu2}) can be done in several different ways. For example, one can construct a formula in the style of bounded model checking. Here we apply \emph{symbolic execution} to generate $F$. Below we first review the standard symbolic execution algorithm and then explain how we use it to construct the formula $F$.

First, we define the set $\Var_0=\{v_0 \mid v \in \Var \}\cup \{c_0\}$ as the set of initial symbolic values of all variables, where $c_0$ is used to denote the symbolic value from the input.
A \emph{symbolic state} is a triple $\langle pc, val, cmds \rangle$, where the \emph{path condition} $pc$ is a guard in $\Grd$ over $\Var_0$, the \emph{symbolic valuation} is a mapping from variables in $\Var$ to an expression in $\Exp$ over $\Var_0$, and $cmds\in \Cmd$ is the remaining commands to be executed. The \emph{initial symbolic state} is $\langle \true, val_0, s_2\rangle$, where $val_0(v) = v_0$ for all $v\in\Var$ and $s_2$ is the loop body of the reducer.
One can compute the next symbolic state of $\langle pc, val, c;cmds \rangle$ according to the following transition rules. Recall that $c$ is either in the form of $v:= e$ or $\ite{g}{c_1}{c_2}$ after the program transformation tasks are performed:
\begin{enumerate}
	\item If $c$ is $v:= e$, the next symbolic state is $\langle pc, val[v \rightarrow val(e) ], cmds \rangle$.
	\item If $c$ is $\ite{g}{c_1}{c_2}$, the next symbolic states are $\langle pc\wedge g, val, c_1.cmds \rangle$ and $\langle pc\wedge not\  g, val, c_2.cmds \rangle$.
\end{enumerate}

Here we lift the $val$ function from variables to expressions in the standard way and $val[v \rightarrow val(e) ]$ is obtained by reassigning the value of $v$ in $val$ to $val(e)$. The execution from a symbolic state terminates when its path condition is conflict or the commands to be executed become empty, i.e., all commands have been processed. So far it is standard.

For the loop body $s_2$, symbolic execution is guaranteed to terminate because there exist no commands in $s_2$ that may lead to infinite computations.
When the execution terminates, we collect all states with empty commands $\langle pc, val, \emptyset \rangle$ as a set $T$ and use them to create the function $F$ in Equation~(\ref{eq:commu2}).
First, we create for each variable $x\in\Var$ a formula $f_x$ whose value is $\bigvee_{\langle pc, val, \emptyset \rangle \in T} (pc \wedge val(x))$.
The function $F(n,x_1,x_2,\ldots,x_k)$ simply returns the tuple $(f_{x_1},f_{x_2},\ldots, f_{x_k})$.

However, doing symbolic execution in a naive way can be extremely inefficient.
The length of the formula $val(v)$ can grow exponentially w.r.t. the input commands. For example, consider a simple command $u:=x+y;u:=u+u;u:=u+u$, each execution of $u:= u+ u$ will double the length of the expression representing the symbolic value of $u$. More concretely, after the execution of $u:=x+y$, the symbolic value of $u$ is $x_0+y_0$. After the two executions of $u:=u+u$, the value of $u$ changes to $x_0+y_0+x_0+y_0$ and $x_0+y_0+x_0+y_0+x_0+y_0+x_0+y_0$, respectively. One way to alleviate the problem is to apply formula simplification algorithms, e.g., those implemented in Z3.
Here we propose in a very simple alternative solution that works quite well in our reducer benchmarks.

We use a more efficient data structure to store symbolic values. At a high level, in a symbolic state, we count the number of occurrences of each symbolic values in $\Var_0$ and update the count when operations such as $+$ and $-$ is applied. For other operations $x\odot y$, we create a new symbolic expression $val(x) \odot val(y)$ and count its number of occurrence as one.
More concretely, we use a map $sv_x:\Exp\rightarrow \mathbb{N}$ (symbolic value of $x$) to remember the number of occurrences of each expression in $x$. We have $val(x)= (\Sigma_{z\in \dom{sv_x}} (z\times sv_x(z)) )$, where $\dom{sv}=\{e\mid sv(e)\neq 0\}$.
Initially, $sv_x(x_0) = 1$ and $sv_x (e) =0$ for all $e\neq x$. Assume that the symbolic value of $x,y$ are $sv_x,sv_y$ in the current symbolic state. The symbolic value $sv'_x$ of $x$ after the assignment $x:=p\circ q$ is defined as follows.

\begin{itemize}
	\item For the case that $\circ \in\{+,-\}$, we have $sv'_x(z) = sv_p(z)\circ  sv_q(z)$.
	\item For the case that $\circ \notin\{+,-\}$, we define $sv'_x(val(p) \circ val(q)) =1$ and $sv'_x(e) = 0$ for all $e\neq val(p) \circ val(q)$.
\end{itemize}
To make the definition complete for our simple reducer program commands, we further define (I) $sv_{\cur}(c_0) =1$ and $sv_{\cur}(e)=0$ for $e\neq \cur$, (II) $sv_{n}(1) =n$ and $sv_{n}(e)=0$ for $e\neq 1 \wedge n\in \Z$.
This data structure is efficient for strong symbolic values of reducer programs mainly for reason that the operations $+$ and $-$ are quite heavily used in such programs.
Here we give a tiny example. Initially, $sv_x(x_0)=1$. After the commands $x:=x*x;x:=x+x$, we get $sv_x(x_0*x_0)=1$ and $sv_x(x_0*x_0)=2$, respectively.}

\section{Evaluation}
\label{section:exp}
J-ReCoVer is implemented in Java and built on top of Soot 2.5.0 and Z3 4.7.1. We run J-ReCoVer on a virtual machine running on a server with AMD Opteron 6376 CPU. 4GB of memory is allocated to this virtual machine. The operation system is Ubuntu 16.04.5 LTS.

In order to properly evaluate the performance of J-ReCoVer, we collect benchmark example from two different sources.
First, we use the search engine \emph{searchcode.com} to collect java programs containing the key strings ``public void reduce'' or ``protected void reduce''. Since there is an upper bound on the number of results returned from the search engine, we add different search filters in order to get more data. We tried all $12$ combinations of six filters on code length $\{<50$, $ 50\sim 250$, $250\sim 450$, $450\sim 650$, $650\sim 850$, $850\sim 1050$, $1050\sim 1250\}$
and two filters for data sources $\{\url{github.com}, \url{bitbuckect.com}\}$. In total, we get $11346$ Java programs. We exclude cases that are not Hadoop MapReduce reducer programs (those do not import the Hadoop library, do not extend or implement the reducer interface) and obtain $1273$ examples. We further remove those that cannot be compiled, those that are identical to each other, or those with non-numerical data types (e.g, string) and obtained $118$ programs as the final benchmarks.

\hide{The second set of examples are those have been discussed in verification literature~\cite{ChenHSW15,ChenSW16}. This set is relatively small, we have only $10$ examples.
We also created a program that randomly generates reducer programs in order to test the scalability of J-ReCoVer.
All benchmark examples can be found on the website of J-ReCoVer.}

For the optimization based on live variable analysis (Section~\ref{section:optimizations}), we evaluate them based on real-world benchmark sets collected from open repositories (118 examples). The difference between the version with and without this optimization is very clear. The results are presented in table~\ref{tab:opt1}. The first column is the result without optimization and the second column is the result with optimization. The precision of the analysis is improved from $6.8\%$ to $97.5\%$. There are only three cases that J-ReCoVer with data flow analysis cannot handle. The reason that J-ReCoVer fails is that the three reducers use more complicated structure than what J-ReCoVer currently supports; they use a branch statement before entering the loop, i.e., in the form $\ite{g}{(s_1;\rloop\{s_2\};s_3)}{(s'_1;\rloop\{s'_2\};s'_3)}$. In theory, such a program can be handled by more sophisticated program transformation tasks. For example, we can merge the two loops and push the outer branch condition into the merged loop.
However, this new program transformation task is not implemented in J-ReCoVer yet.

If we look into scrutiny, we failed to verify any example using Equation~(\ref{eq:commu}). The reason is that when Soot converts a Java program to the Jimple format, it always uses a temporary variable to store the value of $\cur$ in the loop body (similar to the example in Figure~\ref{fig:reducer_opt}). We tried a simpler heuristics approach that only ignores the temporary variables and the result is not as good as using live variable analysis, because the values of the temporary variables often also propagate to other variables.
\begin{table}[htb]
	\centering
	\begin{tabular}{|l|l|l|}
		\hline
		&W/O optimization	& With optimization\\
		\hline
		\hline
		Commutative& 0&106\\
		\hline
		Non-commutative&8&9\\
		\hline
		Unknown&110&3\\
		\hline
		Precision& 6.8\% & 97.5\%\\
		\hline

	\end{tabular}
	\caption{The improvement of precision using data flow analysis. The row ``Precision'' is calculated as the number of examples that J-ReCoVer returns commutative or non-commutative divided by the total number of example.}
	\label{tab:opt1}
\end{table}

\bibliographystyle{splncs-sorted}
\bibliography{refs}

\end{document}