### 2019-06-25 appendix.tex
polish appendix

% \input{appendix}
\clearpage
\appendix

\section{J-ReCoVer Demo}

Our plan is to demo the basic usage of J-ReCoVer over two examples. The first example is a reducer that is not commutative, and J-ReCoVer will report a counterexample. Based on this counterexample, we will fix the reducer to get the 2nd example, and let J-ReCoVer to check it again. Now J-ReCoVer will report the reducer is commutative. Below we give a more detailed description of the functionalities of J-ReCoVer.

\subsection{The User Interface of J-ReCoVer}
\label{appendix:1}

Figure~\ref{fig:main_screen} shows the main screen of J-ReCoVer. The alignment of user interface elements are as follows:

\begin{figure}
\begin{center}
% \vspace{-10pt}
\includegraphics[width=.9\linewidth]{screenshots/main_screen}
% \vspace{-5pt}
\caption{Main Screen of J-ReCoVer}
\label{fig:main_screen}
% \vspace{-15pt}
\end{center}
\end{figure}

\begin{itemize}
\item
Top Menu:
\begin{itemize}
\item
\emph{``GitHub''} opens a new tab of the repository of J-ReCoVer. The repository also has source code of benchmarks.
\item
\emph{``Benchmark''} leads to the list of benchmarks we collected with their results of commutativity analysis as Fig.~\ref{fig:benchmark_list} shows. Select the filename or "try it" button will select the example and go back to the main screen.
\end{itemize}
\item
Right Menu:
\begin{itemize}
\item
\emph{``Benchmark List''} provides a quick selection to examples in the benchmark (Fig.~\ref{fig:benchmark_list_right})
\item
\emph{``Results''} shows the last analysis result.
\end{itemize}
\item
Upper-Right Menu:
\begin{itemize}
\item
\emph{``Type''} provides types that J-ReCoVer supports. Details are explained in~\ref{appendix:3}.
\item
\emph{``Run''} starts the analysis of the selected example.
\end{itemize}
\item
Upper-Left Menu: See~\ref{appendix:3}.
\end{itemize}

\begin{figure}
\begin{center}
% \vspace{-10pt}
\includegraphics[width=.8\linewidth]{screenshots/benchmark_list}
% \vspace{-5pt}
\caption{Benchmark list from top menu (part)}
\label{fig:benchmark_list}
% \vspace{-15pt}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
% \vspace{-10pt}
\includegraphics[width=.8\linewidth]{screenshots/benchmark_list_right}
% \vspace{-5pt}
\caption{Benchmark list from right menu (part)}
\label{fig:benchmark_list_right}
% \vspace{-15pt}
\end{center}
\end{figure}

\subsection{Run analysis of selected benchmark example}
\label{appendix:2}

To run analysis, you can choose an example from benchmark using either the top menu or the right menu, then click the "Run" button of the upper-right menu to start the commutativity analysis of the selected example. The pull-down menu "Types" of the upper-right menu provides selectable types for input \texttt{(T1,T2)} and output \texttt{(T3,T4)} key-value pairs. Fig.~\ref{fig:type_select_menu} shows the pull-down menu and Fig.~\ref{fig:selectable_types} shows part of the selectable types in J-ReCoVer. If one selects an example from the benchmark, the types will be automatically selected to match the example. The selection will be needed if the user wants to write a reduce function from sketch for analysis (see~\ref{appendix:3}).

\begin{figure}
\begin{center}
% \vspace{-10pt}
\includegraphics[width=.8\linewidth]{screenshots/type_select_menu}
% \vspace{-5pt}
\caption{Type select menu}
\label{fig:type_select_menu}
% \vspace{-15pt}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
% \vspace{-10pt}
\includegraphics[width=.8\linewidth]{screenshots/selectable_types}
% \vspace{-5pt}
\caption{Selectable types (part)}
\label{fig:selectable_types}
% \vspace{-15pt}
\end{center}
\end{figure}

After J-ReCoVer finishes the analysis, the \emph{"Result:"} tab of the right menu will show the result of the analysis, which can be found in Figure~\ref{fig:analysis_result}. Figure~\ref{fig:analysis_result_text} shows the complete analysis report of the benchmark \texttt{dis.java} in Figure~\ref{fig:main_screen}.

\begin{figure}
\begin{center}
% \vspace{-10pt}
\includegraphics[width=.8\linewidth]{screenshots/analysis_result}
% \vspace{-5pt}
\caption{The analysis result}
\label{fig:analysis_result}
% \vspace{-15pt}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\begin{mdframed}[roundcorner=5pt]
\begin{Verbatim}[fontsize=\tiny]
================================
### Compile ###
================================
Success

================================
### j-ReCoVer ###
================================
RESULT: Proved to be commutative

________________________________

Trying to find a counterexample...

RESULT: Cannot find a counterexample in 1500 tests
Testcases were generated randomly.

Following are the first 50 lines of testcases:
Input1: o [5, 4, 5, 2, -1]
Input2: o [-1, 4, 5, 5, 2]
Output1:

Key: [o]
Val: [0.0]
--
Output2:

Key: [o]
Val: [0.0]
--------
Input1: j [1, -1, 0, 3, -2]
Input2: j [0, 3, -2, 1, -1]
Output1:

Key: [j]
Val: [0.0]
--
Output2:

Key: [j]
Val: [0.0]
--------
Input1: e [4, 3, 5, -4, 1]
Input2: e [4, -4, 5, 3, 1]
Output1:

Key: [e]
Val: [0.0]
--
Output2:

Key: [e]
Val: [0.0]
--------
Input1: s [3, 0, 2, -4, -1]
Input2: s [3, -1, 0, -4, 2]
Output1:

Key: [s]
Val: [0.0]
--
Output2:

Key: [s]
Val: [0.0]
--------
Input1: v [-3, -2, 3, 3, -3]
Input2: v [3, -2, -3, -3, 3]
\end{Verbatim}
\end{mdframed}
\caption{The analysis result of \texttt{dis.java}}
\label{fig:analysis_result_text}
\end{center}
\end{figure}

In a commutativity analysis, J-ReCoVer first tries to compile the example then perform commutativity analysis. If the analysis can not conclude that the example is commutative, J-ReCoVer then runs 1500 randomly generated test cases to try to find a counterexample. If the test cases all pass, then the result will be "unknown".

\subsection{Run analysis of an example from sketch}
\label{appendix:3}

Besides examples from the benchmark, J-ReCoVer also provides an interactive way to analyse a user defined reducer function.
The upper-left menu (below the top menu) provides blank templates for two versions of Hadoop. The templates are

\begin{mdframed}[roundcorner=5pt]
\begin{verbatim}
public void reduce(T1 key, Iterator<T2> values,
	OutputCollector<T3,T4> oc1, Reporter reporter)
		throws IOException,InterruptedException {

}
\end{verbatim}
\end{mdframed}

, and

\begin{mdframed}[roundcorner=5pt]
\begin{verbatim}
public void reduce(T1 key, Iterable<T2> values,
	Context context) throws IOException,InterruptedException {

}
\end{verbatim}
\end{mdframed}

In this case, one would have to manually fill the types T1 to T4 and the function body, then selects the types from the pull-down menu, as shown in Fig.~\ref{fig:type_select_menu} and Fig.~\ref{fig:selectable_types}.

For more details, all selectable types are:

\begin{itemize}
\item Long
\item BooleanWritable
\item BytesWritable
\item Text
\item Double
\item DoubleWritable
\item Float
\item FloatWritable
\end{itemize}